---
title: "What Can Survey Data Tell Us About Ideological Differences Between Black Voters and Black Nonvoters?"
output:
  md_document:
    variant: gfm+footnotes
    preserve_yaml: TRUE
    pandoc_args: 
      - "--wrap=preserve"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../../_posts") })
date: 2021-11-23
permalink: /posts/2022/07/black-voters-and-nonvoters
toc: true
tags:
  - CES
  - black vote
  - race and ethnicity
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```



```{r}
library(here)
library(tidyverse)
library(survey)
library(ggalt)
```

```{r}
ces <- read.csv("../../../Sandbox/voter-nonvoters2020/ces20/CES20_Common_OUTPUT_vv.csv")
black <- subset(ces, race==2)
black <- black %>%
  filter(!is.na(commonpostweight)) %>%
  mutate(voter = case_when(
    is.na(CL_2020gvm) == FALSE ~ 1,
    is.na(CL_2020gvm) == TRUE ~ 0
  ),
  college = case_when(educ < 3 ~ "No college",
                     educ > 3 ~ "College"),
  gmw = case_when(CC20_350b == 1 ~ 1,
                 CC20_350b == 2 ~ 0),
  oequalpay = case_when(CC20_350d == 1 ~ 1,
                 CC20_350d == 2 ~ 0),
  am4a = case_when(CC20_327a == 1 ~ 1,
                 CC20_327a == 2 ~ 0),
  lowerage = case_when(CC20_327c == 1 ~ 1,
                 CC20_327c == 2 ~ 0),
  zaca = case_when(CC20_327d == 1 ~ 1,
                 CC20_327d == 2 ~ 0),
  tmandate = case_when(CC20_327e == 1 ~ 1,
                 CC20_327e == 2 ~ 0),
  yarban = case_when(CC20_330b == 1 ~ 1,
                 CC20_330b == 2 ~ 0),
  sconceal = case_when(CC20_330c == 1 ~ 1,
                 CC20_330c == 2 ~ 0),
  bpatrols = case_when(CC20_331b == 1 ~ 1,
                 CC20_331b == 2 ~ 0),
  wdeptreports = case_when(CC20_331c == 1 ~ 1,
                 CC20_331c == 2 ~ 0),
  pspendwall = case_when(CC20_331e == 1 ~ 1,
                 CC20_331e == 2 ~ 0),
  nchoice = case_when(CC20_332a == 1 ~ 1,
                 CC20_332a == 2 ~ 0),
  jabortsafe = case_when(CC20_332b == 1 ~ 1,
                 CC20_332b == 2 ~ 0),
  kabortban = case_when(CC20_332f == 1 ~ 1,
                 CC20_332f == 2 ~ 0),
  eweek20 = case_when(CC20_332c == 1 ~ 1,
                 CC20_332c == 2 ~ 0),
  vmnsnt= case_when(CC20_334a == 1 ~ 1,
                 CC20_334a == 2 ~ 0),
  rbodycam = case_when(CC20_334b == 1 ~ 1,
                 CC20_334b == 2 ~ 0),
  dfund = case_when(CC20_334c == 1 ~ 1,
                 CC20_334c == 2 ~ 0),
  cdefund = case_when(CC20_334d == 1 ~ 1,
                 CC20_334d == 2 ~ 0),
  uchoke = case_when(CC20_334e == 1 ~ 1,
                 CC20_334e == 2 ~ 0),
  paris = case_when(CC20_355a == 1 ~ 1,
                 CC20_355a == 2 ~ 0),
  ltransban = case_when(CC20_355d == 1 ~ 1,
                 CC20_355d == 2 ~ 0),
  fsnap = case_when(CC20_355e == 1 ~ 1,
                 CC20_355e == 2 ~ 0),
  mwadv = case_when(CC20_440a < 3 ~ 1, # Agree
                     CC20_440a == 3 ~ 0, # Neither Agree nor Disagree
                     CC20_440a > 3 ~ 0), # Disagree 
  saisolate = # Racial problems in the U.S. are rare, saisolate situations.
           case_when(CC20_440b < 3 ~ 1, # Agree
                     CC20_440b == 3 ~ 0, # Neither Agree nor Disagree
                     CC20_440b > 3 ~ 0), # Disagree
  qwadv2 = # Structural racism
           case_when(CC20_441f < 3 ~ 1, # Agree
                     CC20_441f == 3 ~ 0, # Neither Agree nor Disagree
                     CC20_441f > 3 ~ 0), # Disagree 
  sexism_b = # Women are too easily offended
           case_when(CC20_440b < 3 ~ 1, # Agree
                     CC20_440b == 3 ~ 0, # Neither Agree nor Disagree
                     CC20_440b > 3 ~ 0), # Disagree 
  istructural = # Structural racism
           case_when(CC20_441b < 3 ~ 1, # Agree
                     CC20_441b == 3 ~ 0, # Neither Agree nor Disagree
                     CC20_441b > 3 ~ 0), # Disagree 
  hpolice.unsafe = case_when(CC20_307 < 3 ~ 0,
                     CC20_307 > 2 ~ 1))


blackdat <- svydesign(ids = ~0, 
                    data = black, weights = ~commonpostweight)

```



```{r}
voted <- svytable(~voter, design=blackdat) %>%
  prop.table() %>%
  as.data.frame() %>%
  mutate(Freq = round(Freq * 100, digits = 0))

# gmw, fsnap, oequalpay ####
 
gmw <- svyby(~gmw, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = gmw + 1.96 * se,
         lower = gmw - 1.96 * se)


oequalpay <- svyby(~oequalpay, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = oequalpay + 1.96 * se,
         lower = oequalpay - 1.96 * se) # significant

fsnap <- svyby(~fsnap, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = fsnap + 1.96 * se,
         lower = fsnap - 1.96 * se)

# am4a, aca, lowerage, tmandate ####
am4a <- svyby(~am4a, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = am4a + 1.96 * se,
         lower = am4a - 1.96 * se)

zaca <- svyby(~zaca, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = zaca + 1.96 * se,
         lower = zaca - 1.96 * se) # significant!

lowerage <- svyby(~lowerage, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = lowerage + 1.96 * se,
         lower = lowerage - 1.96 * se)

tmandate <- svyby(~tmandate, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = tmandate + 1.96 * se,
         lower = tmandate - 1.96 * se) # significant!

# arban, sconceal ####
yarban <- svyby(~yarban, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = yarban + 1.96 * se,
         lower = yarban - 1.96 * se) # significant!

sconceal <- svyby(~sconceal, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = sconceal + 1.96 * se,
         lower = sconceal - 1.96 * se)# significant!

# bpatrols, wdeptreports, pspendwall####
bpatrols <- svyby(~bpatrols, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = bpatrols + 1.96 * se,
         lower = bpatrols - 1.96 * se)

wdeptreports <- svyby(~wdeptreports, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = wdeptreports + 1.96 * se,
         lower = wdeptreports - 1.96 * se)# significant!

pspendwall <- svyby(~pspendwall, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = pspendwall + 1.96 * se,
         lower = pspendwall - 1.96 * se) # significant!

# nchoice, jabortsafe, kabortban, eweek20 ####

nchoice <- svyby(~nchoice, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = nchoice + 1.96 * se,
         lower = nchoice - 1.96 * se)

jabortsafe <- svyby(~jabortsafe, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = jabortsafe + 1.96 * se,
         lower = jabortsafe - 1.96 * se)

kabortban <- svyby(~kabortban, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = kabortban + 1.96 * se,
         lower = kabortban - 1.96 * se)

eweek20 <- svyby(~eweek20, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = eweek20 + 1.96 * se,
         lower = eweek20 - 1.96 * se)

# vmnsnt, rbodycam, fund, cdefund, uchoke, hpolice.unsafe ####

vmnsnt <- svyby(~vmnsnt, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = vmnsnt + 1.96 * se,
         lower = vmnsnt - 1.96 * se) # significant

rbodycam <- svyby(~rbodycam, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = rbodycam + 1.96 * se,
         lower = rbodycam - 1.96 * se) # significant

dfund <- svyby(~dfund, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = dfund + 1.96 * se,
         lower = dfund - 1.96 * se)

cdefund <- svyby(~cdefund, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = cdefund + 1.96 * se,
         lower = cdefund - 1.96 * se)

uchoke <- svyby(~uchoke, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = uchoke + 1.96 * se,
         lower = uchoke - 1.96 * se) # significant

hpolice.unsafe <- svyby(~hpolice.unsafe, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = hpolice.unsafe + 1.96 * se,
         lower = hpolice.unsafe - 1.96 * se)

# ltransban, mwadv, saisolate, sexism_b ####
ltransban <- svyby(~ltransban, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = ltransban + 1.96 * se,
         lower = ltransban - 1.96 * se)

mwadv <- svyby(~mwadv, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = mwadv + 1.96 * se,
         lower = mwadv - 1.96 * se) #uhhhh just barely significant

qwadv2 <- svyby(~qwadv2, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = qwadv2 + 1.96 * se,
         lower = qwadv2 - 1.96 * se) # significant

saisolate <- svyby(~saisolate, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = saisolate + 1.96 * se,
         lower = saisolate - 1.96 * se) # significant

sexism_b <- svyby(~sexism_b, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = sexism_b + 1.96 * se,
         lower = sexism_b - 1.96 * se) # significant

istructural <- svyby(~istructural, ~voter, blackdat, svymean, na.rm = TRUE) %>%
  mutate(upper = istructural + 1.96 * se,
         lower = istructural - 1.96 * se)

med_wide <- cbind(am4a, zaca, lowerage, tmandate) %>%
  mutate(type = c("Nonvoter", "Voter")) %>%
  select(-contains("se") & 
             -contains("voter") &
             -contains("upper") &
             -contains("lower")) %>%
  pivot_longer(cols = c(am4a:tmandate),
               names_to = "issue", 
               values_to = "value") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(Nonvoter = Nonvoter * 100,
         Voter = Voter * 100,
         difference = Nonvoter - Voter)

gun_wide <-  cbind(yarban, sconceal) %>%
  mutate(type = c("Nonvoter", "Voter")) %>%
  select(-contains("se") & 
             -contains("voter") &
             -contains("upper") &
             -contains("lower")) %>%
  pivot_longer(cols = c(yarban:sconceal),
               names_to = "issue", 
               values_to = "value") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(Nonvoter = Nonvoter * 100,
         Voter = Voter * 100,
         difference = Nonvoter - Voter)

imm_wide <-  cbind(bpatrols, wdeptreports, pspendwall) %>%    
  mutate(type = c("Nonvoter", "Voter")) %>%
  select(-contains("se") & 
             -contains("voter") &
             -contains("upper") &
             -contains("lower")) %>%
  pivot_longer(cols = c(bpatrols:pspendwall),
               names_to = "issue", 
               values_to = "value") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(Nonvoter = Nonvoter * 100,
         Voter = Voter * 100,
         difference = Nonvoter - Voter)

labor_wide <- cbind(gmw, fsnap, oequalpay) %>%    
  mutate(type = c("Nonvoter", "Voter")) %>%
  select(-contains("se") & 
             -contains("voter") &
             -contains("upper") &
             -contains("lower")) %>%
  pivot_longer(cols = c(gmw:oequalpay),
               names_to = "issue", 
               values_to = "value") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(Nonvoter = Nonvoter * 100,
         Voter = Voter * 100,
         difference = Nonvoter - Voter)

abort_wide <- cbind(nchoice, jabortsafe, kabortban, eweek20) %>%    
  mutate(type = c("Nonvoter", "Voter")) %>%
  select(-contains("se") & 
             -contains("voter") &
             -contains("upper") &
             -contains("lower")) %>%
  pivot_longer(cols = c(nchoice:eweek20),
               names_to = "issue", 
               values_to = "value") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(Nonvoter = Nonvoter * 100,
         Voter = Voter * 100,
         difference = Nonvoter - Voter)

police_wide <- cbind(vmnsnt, rbodycam, dfund, cdefund, uchoke, hpolice.unsafe) %>%    
  mutate(type = c("Nonvoter", "Voter")) %>%
  select(-contains("se") & 
             -contains("voter") &
             -contains("upper") &
             -contains("lower")) %>%
  pivot_longer(cols = c(vmnsnt:hpolice.unsafe),
               names_to = "issue", 
               values_to = "value") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(Nonvoter = Nonvoter * 100,
         Voter = Voter * 100,
         difference = Nonvoter - Voter)

culture_wide <- cbind(istructural, mwadv, qwadv2, saisolate, sexism_b, ltransban) %>%
  mutate(type = c("Nonvoter", "Voter")) %>%
  select(-contains("se") & 
             -contains("voter") &
             -contains("upper") &
             -contains("lower")) %>%
  pivot_longer(cols = c(istructural:ltransban),
               names_to = "issue", 
               values_to = "value") %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(Nonvoter = Nonvoter * 100,
         Voter = Voter * 100,
         difference = Nonvoter - Voter)

fulldat <- rbind(med_wide, gun_wide, imm_wide, labor_wide, abort_wide, police_wide, culture_wide) %>%
  mutate(absdif = abs(difference))

```


```{r}
ggplot() +
  geom_dumbbell(data=fulldat, aes(y=issue, x=`Nonvoter`, xend=Voter),
                size=1.5, color="#DADFF7", size_x=3, size_xend = 3, colour_x = "#88498F", colour_xend = "#73956F") + xlab("Percent") + 
  theme_minimal() + labs(caption = "Plot: Zachary L. Hertz\nData: 2020 CES")

ggsave("plot.png", width = 6, height = 6, units = c("in"), dpi=700)

```

An interesting [recently-published paper](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0268134) by Michael Barber and John Holbein has drawn attention to the well-studied existence of a racial turnout gap: that is, the fact that non-Hispanic white Americans vote at higher rates than racial minorities. 


As the 2022 midterm elections approach, and Democrats brace to defend their control of Congress, pundits have continued to attempt to understand the forces driving the erosion in Democratic support among minority voters in 2020. Conversations about Black mobilization tend to operate around an implicit assumption that Black nonvoters are more conservative than Black voters, and many pundits have chosen to interpret the [REWORD ]erosion of Democratic support among Black voters in 2020 as a sign that "wokism" has driven minority voters away from the Democratic Party. [CITATION] Yet a quantitative assessment of the merits of this argument remains forthcoming.




devoted greater attention to understanding the issue positions of minority voters and nonvoters

