---
title: "How Sensitive are Partisans to Out-Party Cues?"
output:
  pdf_document
header-includes: 
- \definecolor{yourcolourname}{HTML}{800000}
geometry: margin=0.8in
urlcolor: yourcolourname
---

```{r, include=FALSE}
# Bring in data ####
library(stargazer)
library(tidyverse)
library(htmlwidgets)
library(StatMeasures)

dat <- read.csv("../../../Opinion Lab/classpoll_weighted.csv")

dat$minimumwage[dat$minimumwage>50]<- NA
dat$natsec[dat$natsec>2000] <- NA

# Code pid, with leaners
dat$pid2[dat$political_party < 4 | dat$political_party ==6] <- 1
dat$pid2[dat$political_party > 7 | dat$political_party ==5] <- 2

dat$pid_nolean[dat$political_party<3] <- 1
dat$pid_nolean[dat$political_party>8] <- 2

dat$strongpartisan <- 0
dat$strongpartisan[dat$political_party==1 | 
                     dat$political_party ==10] <- 1
dat$ftparty <- NA

# use fts to measure affective polarization via parties
dat <- dat %>%
  mutate(ftparty = case_when(
    pid2==1 ~ ((dat$ft_dem_1-dat$ft_rep_1)/100),
    pid2==2 ~((dat$ft_rep_1-dat$ft_dem_1)/100)
  ))

# use fts to measure affective polarization via candidates
dat <- dat %>%
  mutate(ftcand = case_when(
    pid2==1 ~ ((dat$ft_biden_1-dat$ft_trump_1)/100),
    pid2==2 ~((dat$ft_trump_1-dat$ft_biden_1)/100)
  ))

# Minimum wage ####
# If Dem, in party-position is:
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 1] <- 10
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 2] <- 12.50
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 3] <- 15
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 4] <- 10
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 5] <- 12.50
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 6] <- 15
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 7] <- 10
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 8] <- 15
dat$ipp_mw[dat$pid2==1 & dat$randz1 == 9] <- 15

# If republican, then the same is their out party position
dat$opp_mw[dat$pid2==2 & dat$randz1 == 1] <- 10
dat$opp_mw[dat$pid2==2 & dat$randz1 == 2] <- 12.50
dat$opp_mw[dat$pid2==2 & dat$randz1 == 3] <- 15
dat$opp_mw[dat$pid2==2 & dat$randz1 == 4] <- 10
dat$opp_mw[dat$pid2==2 & dat$randz1 == 5] <- 12.50
dat$opp_mw[dat$pid2==2 & dat$randz1 == 6] <- 15
dat$opp_mw[dat$pid2==2 & dat$randz1 == 7] <- 10
dat$opp_mw[dat$pid2==2 & dat$randz1 == 8] <- 15
dat$opp_mw[dat$pid2==2 & dat$randz1 == 9] <- 15

# If republican, in party position is
dat$ipp_mw[dat$pid2==2 & dat$randz1 < 4] <- 7.25
dat$ipp_mw[dat$pid2==2 & dat$randz1 == 4] <- 10
dat$ipp_mw[dat$pid2==2 & dat$randz1 == 5] <- 10
dat$ipp_mw[dat$pid2==2 & dat$randz1 == 6] <- 10
dat$ipp_mw[dat$pid2==2 & dat$randz1 == 7] <- 12.50
dat$ipp_mw[dat$pid2==2 & dat$randz1 == 8] <- 15
dat$ipp_mw[dat$pid2==2 & dat$randz1 == 9] <- 12.50

# if dem, then same is the out party
dat$opp_mw[dat$pid2==1 & dat$randz1 < 4] <- 7.25
dat$opp_mw[dat$pid2==1 & dat$randz1 == 4] <- 10
dat$opp_mw[dat$pid2==1 & dat$randz1 == 5] <- 10
dat$opp_mw[dat$pid2==1 & dat$randz1 == 6] <- 10
dat$opp_mw[dat$pid2==1 & dat$randz1 == 7] <- 12.50
dat$opp_mw[dat$pid2==1 & dat$randz1 == 8] <- 15
dat$opp_mw[dat$pid2==1 & dat$randz1 == 9] <- 12.50

# Run this when tweaking pid2
dat$party<- NA
dat$party[dat$pid2==1] <- "Democrat"
dat$party[dat$pid2==2] <- "Republican"

dat$party2<- NA
dat$party2[dat$pid_nolean==1] <- "Democrat"
dat$party2[dat$pid_nolean==2] <- "Republican"

dat <- subset(dat, is.na(dat$pid2)==FALSE)
```

```{r minimum wage lm, include=FALSE}
modelmw <- lm(minimumwage ~ ipp_mw + opp_mw + as.factor(party), data=dat)
# rename variables ####
names(modelmw$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'Republican')

modelmwdems <- lm(minimumwage ~ ipp_mw + opp_mw, data=subset(dat, party=="Democrat"))
names(modelmwdems$coefficients) <- c('Intercept','In-Party Position','Out-Party Position')

modelmwreps <- lm(minimumwage ~ ipp_mw + opp_mw, data=subset(dat, party=="Republican"))
names(modelmwreps$coefficients) <- c('Intercept','In-Party Position','Out-Party Position')

modelmw1 <- lm(minimumwage ~ ipp_mw + opp_mw + 
                 strongpartisan + ipp_mw:strongpartisan + 
                 opp_mw:strongpartisan + as.factor(party), data=dat)
names(modelmw1$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'Strong Partisan', 'Republican', 'In-Party Position * Strong Partisan','Out-Party Position * Strong Partisan')

modelmw2 <- lm(minimumwage ~ ipp_mw + opp_mw + 
                 ftparty + ipp_mw:ftparty + 
                 opp_mw:ftparty + as.factor(party), data=dat)
names(modelmw2$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'Affective Polarization (Party)', 'Republican', 'In-Party Position * Affective Polarization (Party)','Out-Party Position * Affective Polarization (Party)')

modelmw3 <- lm(minimumwage ~ ipp_mw + opp_mw + 
                 ftcand + ipp_mw:ftcand + 
                 opp_mw:ftcand + as.factor(party), data=dat)
names(modelmw3$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'Affective Polarization (Candidate)', 'Republican', 'In-Party Position * Affective Polarization (Candidate)','Out-Party Position * Affective Polarization (Candidate)')

```

```{r infrastructure, include=FALSE}
# If Dem, in party-position is:
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 1] <- 1.5
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 2] <- 2
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 3] <- 2.5
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 4] <- 1.5
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 5] <- 2
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 6] <- 2.5
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 7] <- 1.5
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 8] <- 2.5
dat$ipp_inf[dat$pid2==1 & dat$randz3 == 9] <- 2.5

# If republican, then the same is their out party position
dat$opp_inf[dat$pid2==2 & dat$randz3 == 1] <- 1.5
dat$opp_inf[dat$pid2==2 & dat$randz3 == 2] <- 2
dat$opp_inf[dat$pid2==2 & dat$randz3 == 3] <- 2.5
dat$opp_inf[dat$pid2==2 & dat$randz3 == 4] <- 1.5
dat$opp_inf[dat$pid2==2 & dat$randz3 == 5] <- 2
dat$opp_inf[dat$pid2==2 & dat$randz3 == 6] <- 2.5
dat$opp_inf[dat$pid2==2 & dat$randz3 == 7] <- 1.5
dat$opp_inf[dat$pid2==2 & dat$randz3 == 8] <- 2.5
dat$opp_inf[dat$pid2==2 & dat$randz3 == 9] <- 2.5

# If republican, in party position is
dat$ipp_inf[dat$pid2==2 & dat$randz3 < 4] <- 1
dat$ipp_inf[dat$pid2==2 & dat$randz3 == 4] <- 1.5
dat$ipp_inf[dat$pid2==2 & dat$randz3 == 5] <- 1.5
dat$ipp_inf[dat$pid2==2 & dat$randz3 == 6] <- 1.5
dat$ipp_inf[dat$pid2==2 & dat$randz3 == 7] <- 2
dat$ipp_inf[dat$pid2==2 & dat$randz3 == 8] <- 2
dat$ipp_inf[dat$pid2==2 & dat$randz3 == 9] <- 2

# if dem, then same is the out party
dat$opp_inf[dat$pid2==1 & dat$randz3 < 4] <- 1
dat$opp_inf[dat$pid2==1 & dat$randz3 == 4] <- 1.5
dat$opp_inf[dat$pid2==1 & dat$randz3 == 5] <- 1.5
dat$opp_inf[dat$pid2==1 & dat$randz3 == 6] <- 1.5
dat$opp_inf[dat$pid2==1 & dat$randz3 == 7] <- 2
dat$opp_inf[dat$pid2==1 & dat$randz3 == 8] <- 2
dat$opp_inf[dat$pid2==1 & dat$randz3 == 9] <- 2
```

```{r infrastructure lm, include=FALSE}
modelinf <- lm(infrastructure ~ ipp_inf + opp_inf + as.factor(party), data=dat)
names(modelinf$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'Republican')

modelinfdems <- lm(infrastructure ~ ipp_inf + opp_inf, data=subset(dat, party=="Democrat"))
names(modelinfdems$coefficients) <- c('Intercept','In-Party Position','Out-Party Position')

modelinfreps <- lm(infrastructure ~ ipp_inf + opp_inf, data=subset(dat, party=="Republican"))
names(modelinfreps$coefficients) <- c('Intercept','In-Party Position','Out-Party Position')

modelinf1 <- lm(infrastructure ~ ipp_inf + opp_inf + 
                 strongpartisan + ipp_inf:strongpartisan + 
                 opp_inf:strongpartisan + as.factor(party), data=dat)
modelinf2 <- lm(infrastructure ~ ipp_inf + opp_inf + 
                 ftparty + ipp_inf:ftparty + 
                 opp_inf:ftparty + as.factor(party), data=dat)
modelinf3 <- lm(infrastructure ~ ipp_inf + opp_inf + 
                 ftcand + ipp_inf:ftcand + 
                 opp_inf:ftcand + as.factor(party), data=dat)
```

```{r natsec, include=FALSE}
# If Dem, in party-position is:
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 1] <- 700
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 2] <- 750
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 3] <- 800
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 4] <- 700
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 5] <- 750
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 6] <- 800
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 7] <- 700
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 8] <- 750
dat$ipp_ns[dat$pid2==1 & dat$randz4 == 9] <- 800

# If republican, then the same is their out party position
dat$opp_ns[dat$pid2==2 & dat$randz4 == 1] <- 700
dat$opp_ns[dat$pid2==2 & dat$randz4 == 2] <- 750
dat$opp_ns[dat$pid2==2 & dat$randz4 == 3] <- 800
dat$opp_ns[dat$pid2==2 & dat$randz4 == 4] <- 700
dat$opp_ns[dat$pid2==2 & dat$randz4 == 5] <- 750
dat$opp_ns[dat$pid2==2 & dat$randz4 == 6] <- 800
dat$opp_ns[dat$pid2==2 & dat$randz4 == 7] <- 700
dat$opp_ns[dat$pid2==2 & dat$randz4 == 8] <- 750
dat$opp_ns[dat$pid2==2 & dat$randz4 == 9] <- 800

# If republican, in party position is
dat$ipp_ns[dat$pid2==2 & dat$randz4 < 4] <- 750
dat$ipp_ns[dat$pid2==2 & dat$randz4 == 4] <- 800
dat$ipp_ns[dat$pid2==2 & dat$randz4 == 5] <- 800
dat$ipp_ns[dat$pid2==2 & dat$randz4 == 6] <- 800
dat$ipp_ns[dat$pid2==2 & dat$randz4 == 7] <- 850
dat$ipp_ns[dat$pid2==2 & dat$randz4 == 8] <- 850
dat$ipp_ns[dat$pid2==2 & dat$randz4 == 9] <- 850

# if dem, then same is the out party
dat$opp_ns[dat$pid2==1 & dat$randz4 < 4] <- 750
dat$opp_ns[dat$pid2==1 & dat$randz4 == 4] <- 800
dat$opp_ns[dat$pid2==1 & dat$randz4 == 5] <- 800
dat$opp_ns[dat$pid2==1 & dat$randz4 == 6] <- 800
dat$opp_ns[dat$pid2==1 & dat$randz4 == 7] <- 850
dat$opp_ns[dat$pid2==1 & dat$randz4 == 8] <- 850
dat$opp_ns[dat$pid2==1 & dat$randz4 == 9] <- 850
```

```{r natsec lm, include=FALSE}
modelnatsec <- lm(natsec ~ ipp_ns + opp_ns + as.factor(party), data=dat)
names(modelnatsec$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'Republican')

modelnatsecdems <- lm(natsec ~ ipp_ns + opp_ns, data=subset(dat, party=="Democrat"))
names(modelnatsecdems$coefficients) <- c('Intercept','In-Party Position','Out-Party Position')

modelnatsecreps <- lm(natsec ~ ipp_ns + opp_ns, data=subset(dat, party=="Republican"))
names(modelnatsecreps$coefficients) <- c('Intercept','In-Party Position','Out-Party Position')

modelns1 <- lm(natsec ~ ipp_ns + opp_ns + 
                 strongpartisan + ipp_ns:strongpartisan + 
                 opp_ns:strongpartisan + as.factor(party), data=dat)
modelns2 <- lm(natsec ~ ipp_ns + opp_ns + 
                 ftparty + ipp_ns:ftparty + 
                 opp_ns:ftparty + as.factor(party), data=dat)
modelns3 <- lm(natsec ~ ipp_ns + opp_ns + 
                 ftcand + ipp_ns:ftcand + 
                 opp_ns:ftcand + as.factor(party), data=dat)
```

```{r, include=FALSE}
nsQ1 <- quantile(dat$natsec, .25)
nsQ3 <- quantile(dat$natsec, .75)
nsIQR <- IQR(dat$natsec)
no_outliers <- subset(dat, dat$natsec> (nsQ1 - 1.5*nsIQR) & dat$natsec< (nsQ3 + 1.5*nsIQR))

infQ1 <- quantile(dat$infrastructure, .25)
infQ3 <- quantile(dat$infrastructure, .75)
infIQR <- IQR(dat$infrastructure)

infno_outliers <- subset(dat, dat$infrastructure> (infQ1 - 1.5*infIQR) & dat$infrastructure< (infQ3 + 1.5*infIQR))

dat <- subset(dat, is.na(minimumwage)==FALSE)

mwQ1 <- quantile(dat$minimumwage, .25)
mwQ3 <- quantile(dat$minimumwage, .75)
mwIQR <- IQR(dat$minimumwage)
mwno_outliers <- subset(dat, dat$minimumwage> (mwQ1 - 1.5*mwIQR) & dat$minimumwage< (mwQ3 + 1.5*mwIQR))

modelmwmin <- lm(minimumwage ~ ipp_mw + opp_mw + as.factor(party), data=mwno_outliers)

# rename variables ####
names(modelmwmin$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'PID: Republican')

modelnatsecmin <- lm(natsec ~ ipp_ns + opp_ns + as.factor(party), data=no_outliers)
names(modelnatsecmin$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'PID: Republican')

modelinfmin <- lm(infrastructure ~ ipp_inf + opp_inf + as.factor(party), data=infno_outliers)
names(modelinfmin$coefficients) <- c('Intercept','In-Party Position','Out-Party Position', 'PID: Republican')

```


A [recent working paper](https://drive.google.com/file/d/1CL9llsfYWU0egabxQhxYFZn4cO4N4iT8/view) by Anthony Fowler and William Howell finds that partisans update their policy beliefs in response to both in- and out-party elite cues. I found the paper especially notable for a few reasons: its findings are well-supported [yet somewhat incongrous](https://www.cambridge.org/core/journals/british-journal-of-political-science/article/party-cues-in-the-news-democratic-elites-republican-backlash-and-the-dynamics-of-climate-skepticism/9B173B1C904EC70361BBBF51C456A343) with both [previous research](https://onlinelibrary.wiley.com/doi/10.1111/j.1540-5907.2011.00541.x) and the oft-cited[^1] expression that "partisanship is a hell of a drug".[^2] Do partisans update their policy beliefs in response to elite cues from both in-party and out-party leaders? To answer this question, I replicate and extend Fowler and Howell's work over three policy areas: the federal minimum wage, national security budget and infrastructure spending.

Research Design
======

For each policy area, I presented respondents with the current policy.[^3] There were nine treatment arms for each policy area, in which Nancy Pelosi was said to support one of three possible positions and Mitch McConnell was said to hold one position from a set of three overlapping but separate positions. Both the assigned treatment and order of politicians' positions within a treatment were randomized. Respondents were then asked for their preferred policy position, and acceptable responses were restricted to a set of plausible numbers for a given policy position.

Respondents were recruited via Lucid between May 2, 2021 and May 9, 2021; I implemented two pre-treatment attention checks into the survey battery to screen for and remove inattentive respondents. While I collected data from 1424 respondents overall after inattentive respondents were removed, I examine partisans (including leaners)[^4] in this analysis, limiting the sample further to 1210 respondents. Finally, I exclude outliers (policy positions further than 1.5 times the IQR below the first quartile or above the third quartile).

Following Fowler and Howell, I begin my analysis by estimating the effects of party leaders' position on partisans' policy beliefs using the following equation for each policy area: $$ \text{Respondent Position} = \alpha + \beta * \text{in-party position} + \gamma * \text{out-party position} + \delta * \text{Party ID} + \epsilon $$


Here, `in-party position` is the policy said to be preferred by the respondent's copartisan party leader while `out-party position` is the policy stated to be preferred by the other party leader. For example, among respondents in the treatment arm where Mitch McConnell preferred a $12.50 minimum wage and Nancy Pelosi preferred a $15 minimum wage, `in-party position` would equal $15 for a Democrat but $12.50 for a Republican, while `out-party position` would equal $12.50 for a Democrat but $15 for a Republican. `Respondent Partisan Identification` is a dummy variable indicating the respondent's partisan identity.

Using this model, the coefficients for party leaders' positions will tell us both the extent and direction of respondents' sensitivity to elite cues. Positive coefficients show that a respondent has positively updated their preferred policy position to more closely match that of the politician, while a negative coefficient shows that the respondent has moved their preferred policy away from the politician's stated position. The strength of these effects is weak when the coefficient is near 0 and increases as the coefficient approaches 1. 

```{r, echo=FALSE, results='asis'}
stargazer(modelmwmin, modelnatsecmin, modelinfmin, 
          omit.stat = c("rsq", "f", "adj.rsq", "ser"),
          title = "Table 1: Effect of In-Party and Out-Party Elite Cues on Policy Preferences",
          label="tab1",
          dep.var.labels = c("Minimum Wage", "National Security", "Infrastructure"),
          header=FALSE, type='latex', style = "apsr")

```

Results
======
Table 1 presents the results of this experiment. As the in-party leader espouses a federal minimum wage that is one dollar higher, copartisan respondents respond with a preferred federal minimum wage that is 17 cents higher (p<0.001). Similarly, for every additional dollar in the federal minimum wage supported by the out-party leader, raise their preferred minimum wage by nearly 16 cents (p<0.001).

For every $1 billion dollar increase in the in-party leader's preferred national security budget, respondents increase their preferred national security budget by $370 million (p=0.02). However, there is no statistically significant effect of the out-party leader's position on respondents' preferred amount of national security spending, nor a noticeable effect of party identification.

Finally, for every additional $1 trillion in infrastructure spending advocated by the in-party leader, partisan respondents increase their preferred amount of infrastructure spending by $433 billion (p<0.001). Additionally, for every additional $1 trillion in the size of the infrastructure bill advocated by the opposing party leader, partisans increase the size of their preferred infrastructure spending bill by $239 billion (p<0.001).

Conclusion
======
I find strong evidence that partisans adjust their preferred federal minimum wage and preferred amount of infrastructure spending in response to cues from both in-party and out-party leaders. I find that on the issue of national security spending, however, only the in-party leader's position has a statistically significant effect on partisans' policy preference. These findings are notable for a few reasons. First, I find that while partisans respond to the positions taken by their party leaders, they do not do so indiscriminately. On all three issues, the positions taken by in-party leaders are discounted significantly by partisan respondents, suggesting that partisans do not blindly adopt the positions of their party leaders.

Second, these findings strengthen the results presented by [Fowler and Howell](https://drive.google.com/file/d/1CL9llsfYWU0egabxQhxYFZn4cO4N4iT8/view): partisans do not broadly dismiss policy preferences from out-party leaders out-of-hand, but rather reconsider their own positions in light of information from leaders of both parties. It is especially notable that these results hold for the issues of the federal minimum wage and an infrastructure bill on a survey fielded in May 2021, given the potential political salience of these issues. Just two months earlier, all 50 Senate Republicans and seven Democrats [voted against](https://www.nytimes.com/2021/03/05/us/minimum-wage-senate.html) a $15 minimum wage, while President Biden had [just proposed](https://www.vox.com/2021/3/31/22357179/biden-two-trillion-infrastructure-jobs-plan-explained) a federal infrastructure bill.

Political salience might provide one potential explanation for why these results failed to replicate on the issue of national security spending. Perhaps the active political debate on the minimum wage and infrastructure bill encourages partisans to actively consider their own positions as well as those of the legislating parties, while on the less-salient issue of national security spending partisans fall back on signals from their party leader. Alternatively, national security spending might simply be an issue where partisans have less faith in the other party to accurately legislate.

Overall, these findings provide reason for optimism as to the state of political discourse between American partisans. Even as [polarization grows](https://www.pewresearch.org/politics/2014/06/12/political-polarization-in-the-american-public/), partisans appear open to considering positions taken by the other side as well as their own.

[^1]: See, for example, [its frequent use](https://twitter.com/search?q=partisanship%20is%20a%20hell%20of%20a%20drug&src=typed_query) by political scientists on Twitter.

[^2]: One [which seems](https://twitter.com/BrendanNyhan/status/857963843607298049) to have been created and popularized by Brendan Nyhan and Stephen Miller.

[^3]: The infrastructure bill had no status-quo equivalent, so respondents were simply presented with the sentence: "Congress is currently considering an infrastructure bill to create jobs and rebuild national infrastructure."

[^4]: While the analysis of pooled partisans excluding leaners is omitted for brevity, doing so actually strengthens the treatment effects slightly to 0.25 for both in-party and out-party positions in the minimum wage experiment.